Learning Libraries of Subroutines for Neurally--Guided Bayesian Program Learning

### 2018-05-15 nips_2018.tex
rational data

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
We automatically generated 109 string editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (a x) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})
  \\\bottomrule
    \end{tabular}
  \caption{Some string editing learned subroutines }\label{textPrimitives}
  \end{table}

\begin{figure}
  \centering\begin{tabular}{ll|lll}\toprule
    Input&Output&Input 1&Input 2&Output\\\midrule
Temple Annalisa Haven 185&TAH1&Launa&Withers&Launa Withers\\
Lara Gregori Bradford&LGB& Rudolf&Akiyama&Rudolf Akiyama
%DkI$>$DwKI$<$xaPoc& 	DWKI\\
%KfvTh$>$Dbl$<$OIJso 	&DBL
\\\midrule
\multicolumn{2}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
\multicolumn{3}{l}{$f(\text{\code{a}},\text{\code{b}}) = $\code{(}$f_0$\code{ a (cons " " b))}}
\\
\bottomrule
  \end{tabular}

@@

\section{Symbolic Regression Experiments}\label{regressionSection}
\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.4\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(* real (+ x real))}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* x (+ real x))} \\
  $f_2($\code{x}$)\,=\,$\code{(/ (/ x real) x)}
  \\\bottomrule
\end{tabular}
\caption{Some learned subroutines for symbolic regression. The system starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&   \textbf{98\% }&94\%&0.7\%&0\%&2\% \\
  Solve time&  2.7s& 2.8s  &3.6s&--&2.1s\\
  %% MDL (nats)&   9.3 &11.0&2.3&--&4.6

@@

\begin{figure}\centering
\includegraphics[width = 3.5cm]{figures/regression.eps} 
  \includegraphics[width = 3.5cm]{figures/list.eps}
  \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        
  \caption{Learning curves for \system both with (blue) and without (red) the recognition model as the frontier size is varied (solid/dashed/dotted lines).}\label{learningCurves} 
\end{figure}

>>>>

\subsection{String Editing}\label{textSection}
Synthesizing programs that manipulate strings is a classic problem in the
programming languages and AI literatures~\cite{menon2013machine,lau2001programming},
and algorithms that learn string editing programs ship in Microsoft Excel~\cite{gulwani2011automating}.
This prior work presumes a ready-made DSL,
expertly crafted to suit string editing.
We show \system can instead start out with generic Lisp primitives
and recover many of the higher-level building blocks that have made these
other system successful.
An obstacle here, however,
is that
our enumerative search procedure
has no means of generating string constants,
and so we incorporate  string-valued parameters as a primitive,
  defining $\probability[x|p]$ by marginalizing out the
  values of the string via dynamic programming.
  In  Section~\ref{regressionSection},
  we will use a similar trick to synthesize programs containing real numbers using gradient descent.

We automatically generated 109 string editing tasks  %% with 4 input/output examples each
(Fig.~\ref{exampleTextProblem}) and model strings as lists of characters.
At first, \system cannot find any correct programs for most of the tasks.
It assembles a DSL (Tbl.~\ref{textPrimitives}) that lets it rapidly explore the space of programs and find solutions to
all of the tasks.

@@

\begin{table}[t]
  \centering
    \begin{tabular}{l}
      \toprule
      $f_0($\code{a}$,$\code{b}$) \,=\, $\code{(fold a b (lambda (x y) (cons x y))))))}\\
      \hspace{0.5cm}($f_0$: \emph{Appends lists (of characters))}\\
      $f_1($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (cdr (if (= c x) s a))))}\\
        \hspace{0.5cm}($f_1$: \emph{Drop first characters from }\code{s}\emph{ until  }\code{c}\emph{ reached})\\
      $f_2($\code{s}$) \,=\, $\code{(unfold s empty? car (lambda (z) (}$f_1$\code{ z SPACE)))}\\
        \hspace{0.5cm}($f_2$: \emph{Abbreviates a sequence of words})\\
        $f_3($\code{s}$,$\code{c}$) \,=\, $\code{(fold s s (lambda (x a) (if (= c x) nil (cons x a))))}\\
        \hspace{0.5cm}($f_3$: \emph{Take characters from }\code{s}\emph{ until }\code{c}\emph{ reached)}
        %(lambda (lambda (fold $0 $0 (lambda (lambda (if (char-eq? $1 $3) empty (cons $1 $0)))))))
  \\\bottomrule
    \end{tabular}
  \caption{Some string editing learned subroutines }\label{textPrimitives}
  \end{table}

\begin{figure}
  \centering\begin{tabular}{r|r}\toprule
Temple Annalisa Haven 185 $\longrightarrow$ TAH1&Nancy FreeHafer $\longrightarrow$ Dr. Nancy\\
Lara Gregori Bradford $\longrightarrow$ LGB& Andrew Cencici $\longrightarrow$ Dr. Andrew
%DkI$>$DwKI$<$xaPoc& 	DWKI\\
%KfvTh$>$Dbl$<$OIJso 	&DBL
\\\midrule
\multicolumn{1}{l|}{$f(\text{\code{s}}) = $\code{(}$f_2$\code{ s)}}&
\multicolumn{1}{l}{$f(\text{\code{s}}) = $\code{(}$f_0$\code{ "Dr. " (}$f_3$\code{ s " "))}}
\\
\bottomrule
  \end{tabular}

@@

\section{Symbolic Regression: Programs from visual input}\label{regressionSection}
We apply \system
to symbolic regression problems.  Here, the
agent observes points along the curve of a function, and must write a
program that fits those points.  We initially equip our learner with
addition, multiplication, and division, and task it with solving
100 symbolic regression problems, each either a polynomial of degree
1--4 or a rational function.  The recognition model is a
convolutional network that observes an image of the target function's
graph (Fig.~\ref{functions}) -- visually, different kinds of
polynomials and rational functions produce different kinds of graphs,
and so the recognition model can learn to look at a graph and predict
what kind of function best explains it.  A key difficulty, however, is
that these problems are best solved with programs containing real
numbers.  Our solution to this difficulty is to allow the system to
write programs with real-valued parameters, and then fit those
parameters by automatically differentiating through the programs the
system writes and use gradient descent to fit the parameters.
We define the likelihood model, $\probability[x|p]$, by assuming a Gaussian noise model for the input/output examples,
and penalize the use of real-valued parameters using the BIC~\cite{Bishop:2006:PRM:1162264}.

\begin{wrapfigure}{r}{0.3\textwidth} \newcommand{\functionSize}{0.75cm}\vspace{-0.5cm}
  \includegraphics[width = \functionSize]{figures/functions/0.png}
  \includegraphics[width = \functionSize]{figures/functions/6.png}
  \includegraphics[width = \functionSize]{figures/functions/207.png}
  \includegraphics[width = \functionSize]{figures/functions/160.png}
  \includegraphics[width = \functionSize]{figures/functions/317.png}\\  
  \includegraphics[width = \functionSize]{figures/functions/432.png}
  \includegraphics[width = \functionSize]{figures/functions/415.png}
  \includegraphics[width = \functionSize]{figures/functions/426.png}
  \includegraphics[width = \functionSize]{figures/functions/597.png}
  \includegraphics[width = \functionSize]{figures/functions/591.png}
  \caption{Recognition model input for symbolic regression. While the DSL learns subroutines for rational functions \& polynomials, the recognition  model jointly learns to look at a graph of the function (above) and predict which of those subroutines is appropriate for explaining the observation.}\label{functions}
\end{wrapfigure}
%% \begin{align*}
%%   \log \probability[x|p] \triangleq& \sum_{n\leq N}\log
%%   \mathcal{N}(p(i_n,\vec{\mathcal{R}}^*)|o_n) - \frac{D\log N}{2}
%% \end{align*}

@@

\system learns a DSL containing templates for polynomials of different orders, as well as ratios of polynomials (Fig.~\ref{regressionDSL}).
The algorithm also discovers programs that minimize the number of continuous degrees of freedom.
For example, it learns to represent linear functions with the program
\code{(* real (+ x real))}}, which has two continuous degrees of freedom, and represents quartic functions using the invented DSL primitive $f_6$ in Tbl.~\ref{regressionDSL}
which has five continuous parameters.
This phenomenon arises from our Bayesian framing -- both the bias towards shorter programs and the likelihood model's BIC penalty.
\begin{wrapfigure}{l}{0.35\textwidth}\centering
\begin{tabular}{l}
  \toprule
  $f_0($\code{x}$)\,=\,$\code{(+ x real)}\\
  $f_1($\code{x}$)\,=\,$\code{($f_0$ (* real x))} \\
  $f_2($\code{x}$)\,=\,$\code{($f_1$ (* x (}$f_0$\code{ x)))}\\
  \hspace{0.5cm}\emph{($f_2$: quadratics)}\\
  $f_3($\code{x}$)\,=\,$\code{(/ (}$f_2$\code{ x) (}$f_0$\code{ x))}\\
    \hspace{0.5cm}\emph{($f_3$: ratio of polynomials)}
  \\\bottomrule
\end{tabular}
\caption{Some learned subroutines for symbolic regression. System starts with addition, multiplication, division, and real numbers, and learns to build rational functions and polynomials up to 4th order.}\label{regressionDSL}
\end{wrapfigure}

@@

  \\\midrule\multicolumn{6}{c}{\emph{Symbolic Regression}}\\\midrule
  \% solved&   \textbf{84\% }&75\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &31s&55s&29s\\
  %% MDL (nats)&   9.3 &11.0&2.3&--&4.6

@@

\begin{figure}\centering
\includegraphics[width = 3.5cm]{figures/rationalCurve.eps} 
  \includegraphics[width = 3.5cm]{figures/list.eps}
  \includegraphics[width = 3.5cm]{figures/textLearningCurve.eps}        
  \caption{Learning curves for \system both with (blue) and without (red) the recognition model as the frontier size is varied (solid/dashed/dotted lines).}\label{learningCurves} 
\end{figure}
